---
title: "mff-overview"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mff-overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(MFF)
```

## Introduction

The **MFF (Meta Fuzzy Functions)** package provides a flexible framework for combining multiple base prediction models using fuzzy and crisp clustering strategies. The main idea is to construct **cluster-specific ensembles** where each cluster learns its own set of weights for aggregating model predictions.

This approach is particularly useful when different data regimes or latent structures favor different predictive models.

------------------------------------------------------------------------

## Core Concepts

Let $X \in \mathbb{R}^{n \times M}$ denote the prediction matrix of $M$ base models for $n$ observations.\
MFF assigns **cluster-specific weight vectors** to aggregate these predictions.

For a given cluster $k$, the predicted values are obtained as:

$$
\mathrm{MFF}_{cluster} = w \times X
$$

where $w_k$ is the weight vector associated with cluster $k$.

------------------------------------------------------------------------

## Training Base Models

Before applying MFF, base models must be trained and their predictions collected into a prediction matrix.

The helper function `model.train()` provides a convenient way to train multiple models and extract their prediction matrices.

```{r}
data <- MASS::Boston

set.seed(123)
result <- model.train(
  target = "medv",
  data = data,
  ntest = 50,
  nvalid = 50
)

str(result$pred_matrix_valid)
```

The resulting prediction matrix has one column per base model and one row per observation.

------------------------------------------------------------------------

## Fitting an MFF Model

### Standard MFF

A standard MFF model learns a **single set of weights** (or cluster-specific weights if clustering is applied).

```{r}
set.seed(123)
mff_model <- mff(x = result$pred_matrix_valid,
y = result$y_valid, c = 6, m = 3)
mff_model
```

------------------------------------------------------------------------

### Tuned MFF

The `tune.mff()` function evaluates multiple MFF configurations and selects the best-performing model according to a chosen metric. With same seed and parameters `mff()` gives
same result with tuning function.

```{r}
mff_tuned <- tune.mff(x = result$pred_matrix_valid,
y = result$y_valid, max_c = 6, iter.max = 100, nstart = 100, mff.method = "fcm",seed = 123)
mff_tuned
```

```{r}
best_weight_sorted <- sort(mff_tuned$best_weight, decreasing = F)

barplot(
  best_weight_sorted,
  horiz = TRUE,
  las = 1,
  col = "steelblue",
  xlab = "Model Weight",
  main = "Optimized Model Weights"
)
```


------------------------------------------------------------------------

## Prediction

Prediction is performed using the generic `predict()` function. The same interface applies to both standard and tuned MFF models.

```{r}
X_test <- result$pred_matrix_test
out <- predict(mff_tuned, X_test)
```

By default, the **best-performing MFF model** is used when available. If the model is not tuned, predictions are generated using the learned weights.

The returned object contains:

-   `mff_preds`: cluster-specific predictions
-   `mff_weights`: corresponding weight matrix (rounded for readability)

```{r}
head(out$mff_preds)
out$mff_weights
```

------------------------------------------------------------------------

## Prediction Modes

The `type` argument controls the prediction mode:

-   `"best"` (default): uses the optimal MFF model if available
-   `"all"`: returns predictions based on all learned cluster configurations

```{r}
out_all <- predict(mff_tuned, X_test, type = "all")
out_all$mff_weights
```

------------------------------------------------------------------------

## Model Evaluation

The `evaluate()` function measures prediction accuracy by comparing model outputs with the observed target values.

It requires two main arguments:

-   `predicted`: the predicted values obtained from an MFF model
-   `actual`: the true observed response values

The function returns standard performance metrics that summarize the predictive quality of the model.

```{r}
evaluate(
  predicted = out$mff_preds,
  actual = result$y_test
)
```

## Summary

The MFF framework enables flexible and interpretable ensemble learning by:

-   combining multiple base models,
-   leveraging fuzzy or crisp clustering,
-   producing cluster-specific predictions,
-   and offering a unified prediction interface.

This design allows users to seamlessly switch between standard and tuned meta-fuzzy models without changing the prediction workflow.
